---
title: "TxDb.Athaliana.Ensembl.TAIR10"
author: "Kabilan"
output:
  BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{TxDb.Athaliana.Ensembl.TAIR10}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

`TxDb.Athaliana.Ensembl.TAIR10` provides transcript annotation
information for *Arabidopsis thaliana* TAIR10.

This vignette demonstrates how to retrieve the transcript database from
`AnnotationHub` and access genes, transcripts, exons, coding sequences,
promoter regions, and feature relationships.

# Installation

Install the package from Bioconductor with:

```{r install, eval=FALSE}
BiocManager::install("TxDb.Athaliana.Ensembl.TAIR10")
```

# Load required packages

```{r loading, message=FALSE, warning=FALSE}
library(AnnotationHub)
library(GenomicFeatures)
library(GenomicRanges)
```

# Retrieve the TxDb object from AnnotationHub

The TxDb object is retrieved using its AnnotationHub record:

```{r annotationhub, message=FALSE, warning=FALSE}
ah <- AnnotationHub::AnnotationHub()
TxDb.Athaliana.Ensembl.TAIR10 <- ah[["AH122287"]]
```

Check the class of the retrieved object:

```{r class}
class(TxDb.Athaliana.Ensembl.TAIR10)
```

# TxDb summary

Display the transcript database:

```{r txdb}
TxDb.Athaliana.Ensembl.TAIR10
```

A summary can also be obtained with:

```{r summary}
summary(TxDb.Athaliana.Ensembl.TAIR10)
```

# Sequence information

The sequence information is available through `seqinfo()`:

```{r seqinfo}
seqinfo(TxDb.Athaliana.Ensembl.TAIR10)
```

The available sequence levels can be obtained with:

```{r seqlevels}
seqlevels(TxDb.Athaliana.Ensembl.TAIR10)
```

# Gene annotations

Gene ranges can be extracted with `genes()`:

```{r genes}
gene_annotations <- genes(TxDb.Athaliana.Ensembl.TAIR10)
gene_annotations
```

# Transcript annotations

Transcript ranges can be extracted with `transcripts()`:

```{r transcripts}
transcript_annotations <- transcripts(TxDb.Athaliana.Ensembl.TAIR10)
transcript_annotations
```

# Exon annotations

Exon ranges can be extracted with `exons()`:

```{r exons}
exon_annotations <- exons(TxDb.Athaliana.Ensembl.TAIR10)
exon_annotations
```

# Coding sequences

Coding sequence ranges can be extracted with `cds()`:

```{r cds}
cds_annotations <- cds(TxDb.Athaliana.Ensembl.TAIR10)
cds_annotations
```

# Promoter regions

Promoter regions can be generated using `promoters()`:

```{r promoters}
promoter_annotations <- promoters(
    TxDb.Athaliana.Ensembl.TAIR10,
    upstream = 2000,
    downstream = 200
)
promoter_annotations
```

# Transcripts grouped by gene

Transcripts can be grouped by gene with `transcriptsBy()`:

```{r transcriptsBy}
transcripts_by_gene <- transcriptsBy(
    TxDb.Athaliana.Ensembl.TAIR10,
    by = "gene"
)
transcripts_by_gene
```

# Exons grouped by transcript

Exons can be grouped by transcript with `exonsBy()`:

```{r exonsBy}
exons_by_transcript <- exonsBy(
    TxDb.Athaliana.Ensembl.TAIR10,
    by = "tx"
)
exons_by_transcript
```

# Gene ranges

The chromosome assignments of genes can be inspected with:

```{r gene-ranges}
head(seqnames(gene_annotations))
```

# Session information

```{r sessioninfo}
sessionInfo()
```
