---
title: "TxDb.Athaliana.Ensembl.TAIR10"
author: "Kabilan"
output: BiocStyle::html_document
vignette: >
  %\VignetteIndexEntry{TxDb.Athaliana.Ensembl.TAIR10}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

# Introduction

**TxDb.Athaliana.Ensembl.TAIR10** is a Bioconductor-compatible
transcript annotation package for *Arabidopsis thaliana* based on the
**TAIR10 reference genome**.

The TxDb object was generated from the **Ensembl Plants GFF3 annotation
(Ensembl Plants release 63)**. The annotation includes genomic
coordinates and transcript models for genes, transcripts, exons, and
coding sequences (CDS).

The package is intended for downstream genomic analyses using
Bioconductor packages such as **GenomicFeatures**, **GenomicRanges**,
**Signac**, **ArchR**, and **ChIPseeker**.

# Loading the package

```{r loading}
library(TxDb.Athaliana.Ensembl.TAIR10)
library(GenomicFeatures)
library(GenomicRanges)
```

# Package summary

The TxDb object can be accessed directly after loading the package.

```{r summary}
TxDb.Athaliana.Ensembl.TAIR10
```

# Genome and chromosome information

The genome information associated with the TxDb can be inspected using
`seqinfo()`.

```{r chromosome}
seqinfo(TxDb.Athaliana.Ensembl.TAIR10)
```

The available chromosome and sequence names can be retrieved using
`seqlevels()`.

```{r avl_chromosomes}
seqlevels(TxDb.Athaliana.Ensembl.TAIR10)
```

The annotation includes the five nuclear chromosomes and the
mitochondrial and plastid sequences of the TAIR10 reference genome.

# Gene annotations

All annotated genes can be retrieved using `genes()`.

```{r gene_annotations}
genes(TxDb.Athaliana.Ensembl.TAIR10)
```

# Transcript annotations

Transcript models can be retrieved using `transcripts()`.

```{r transcripts_annotations}
head(transcripts(TxDb.Athaliana.Ensembl.TAIR10))
```

# Exon annotations

Exon coordinates can be retrieved using `exons()`.

```{r exon_annotations}
head(exons(TxDb.Athaliana.Ensembl.TAIR10))
```

# Coding sequence (CDS) annotations

Coding sequence coordinates can be retrieved using `cds()`.

```{r cds}
head(cds(TxDb.Athaliana.Ensembl.TAIR10))
```

# Promoter regions

Promoter regions can be defined relative to transcription start sites
using the `promoters()` accessor.

For example, the following extracts regions extending 2 kb upstream and
200 bp downstream of transcription start sites.

```{r promoters}
head(promoters(
    TxDb.Athaliana.Ensembl.TAIR10,
    upstream = 2000,
    downstream = 200
))
```

# Transcripts grouped by gene

Transcripts can be grouped according to their parent genes using
`transcriptsBy()`.

```{r transcripts_gene}
tx_by_gene <- GenomicFeatures::transcriptsBy(
    TxDb.Athaliana.Ensembl.TAIR10,
    by = "gene"
)

tx_by_gene
```

# Exons grouped by transcript

Exons can be grouped by transcript using `exonsBy()`.

```{r exons_transcript}
exonsBy(
    TxDb.Athaliana.Ensembl.TAIR10,
    by = "tx"
)
```

# Gene models for downstream analysis

The TxDb can be used as a source of genomic coordinates for downstream
Bioconductor workflows. For example, gene and transcript coordinates can
be converted to `GRanges` objects and used with other genomic datasets.

```{r gene_ranges}
gene_ranges <- genes(TxDb.Athaliana.Ensembl.TAIR10)

gene_ranges
```

# Session information

```{r sessioninfo, eval=TRUE}
sessionInfo()
```
